Efficient FPGA Implementation of\\ Digit Parallel Online Arithmetic Operators

### 2014-05-23 FPT2014/bare_conf.tex
Add extra future works
Davids review

\begin{abstract}
%\boldmath
Online arithmetic has been widely utilized for ASIC implementations with significant performance improvements, as it is designed to perform computations from the most significant digits. Recent research also shows that the digit parallel online operators can fail more gracefully when operating beyond the deterministic region, in comparison to the operators with conventional arithmetic. Unfortunately, the utilization of online arithmetic operators requires large area overhead, and the efficient implementation of general purpose online operators on FPGAs remains under investigation. In this paper, we propose novel approaches to implement key primitives of online arithmetic, such as adders and multipliers, efficiently on modern FPGAs with 6-input LUTs. The FPGA carry resources are used for both area reduction and performance improvements. We demonstrate experimentally that the proposed architectures achieve significant area savings and frequency speed-ups comparing to the original implementations.

@@

\section{Introduction}\label{Sec:Intro}
In the conventional form of computer arithmetic, the computation results are generated either from the least significant digit (LSD), e.g. addition and multiplication, or from the most significant digit (MSD), e.g. division and square root. This inconsistency in computing directions will potentially result in large latency when propagating data among different operations. Online arithmetic was designed to solve this problem \cite{Ercegovac_OnlineOverview,Ercegovac_Book}. With online arithmetic, both inputs and outputs are processed in a MSD-first manner. This enables parallelism and duplication among multiple operations, and the overall computation latency can be significantly reduced. A brief overview of online arithmetic is given in Section~\ref{Sec:Background}.

Recent research has demonstrated another key advantage of online arithmetic that it allows gracefully degradation when operating beyond the deterministic region, e.g. under overclocking \cite{SKDAC14}. This stream of research stems from the inspiration that for certain applications, releasing the absolute accuracy requirements could lead to significant performance improvements. However, traditional form of computer arithmetic does not fail gracefully when timing violation happens, because timing errors initially affect MSDs of the results. In comparison, online arithmetic is more ``overclocking friendly'', as timing errors only occurs at LSDs of the results.

Despite its attractiveness, the usage of online arithmetic on existing FPGAs is still limited due to the huge area overhead. Generally, it is difficult to efficiently map the arithmetic circuits solely onto the look-up-tables (LUTs), which are fundamental building blocks of FPGAs. Lots of efforts has been done to alleviate this concern. For instance, both main commercial FPGA vendors introduced dedicated carry logic into the logic block to reduce the carry propagation delay by over one order of magnitude~\cite{Virtex6}. There is a large amount of studies exploring methods that can effectively map circuits using the carry logic for performance improvements~\cite{FPL10_FPGA_CarryChain,Prefix_CarryChain}, or propose alternative carry logic for FPGAs~\cite{FPGA_CarryChain_New1,FPGA_CarryChain_New2} . Besides, hard DSP blocks are also included to boost multiplications and multiply-accumulations~\cite{XilinxDSP}. However, most of these approaches are designed to accelerate computations with the conventional form of arithmetic.

For online arithmetic, we notice from existing literatures that there are 2 directions that are worth investigating. First, although there has been previous research about FPGA implementations of online arithmetic, they typically only focus on online adders. A brief review of relevant works in this area is presented in Section~\ref{Sec:OA_FPGA}. However, the implementation of other key arithmetic primitives, such as online multipliers, are lack of exploration. Second, we found that existing approaches of building general purpose online adders are specifically for low cost FPGAs with 4-input LUTs and 2 LUTs within a Slice, such as the Altera Cyclone III \cite{AlteraCyclone} and Xilinx Spartan 3 series \cite{XilinxSpartan}. However, as will be discussed in Section~\ref{Sec:OA_FPGA}, these approaches cannot be directly applied on many modern FPGAs with 6-input LUTs and 4 LUTs in a slice, such as the Xilinx Virtex series FPGAs and all Xilinx 7 series FPGAs~\cite{Virtex7}.

In this paper, both issues will be addressed. In Section~\ref{Sec:OA_FPGA}, we propose a novel approach of mapping the digit parallel online adder onto FPGA with 6-LUTs. The mapping is based on the fast carry logic, and the available resources within a slice can be fully utilized. In Section~\ref{Sec:OM_FPGA}, we optimize the online multiplication algorithm to yield an efficient FPGA implementation of digit parallel online multiplier. In addition, the theoretical resource usage of both operators is given, and they are verified experimentally on Xilinx Virtex-6 FPGAs. We demonstrate that the proposed designs achieve significant area reduction and performance gain over their original implementations. In summary, the main contributions of this paper are as follows:
\begin{enumerate}
    \item An efficient approach of mapping online adders onto modern FPGAs;
    \item Area efficient implementation of digit parallel online multiplication algorithm on FPGAs;
    \item Theoretical area models of the proposed architectures.
\end{enumerate}

@@

Online arithmetic has been widely used in numerous applications such as signal processing and control algorithms~\cite{Online_FPGADSP,Online_Control}. Online arithmetic was originally designed for digit-serial operation, of which the data flow is illustrated in Fig~\ref{Fig:OnlineDataFlow}. It can be seen that in order to generate the first output digit, $\delta$ digits of inputs are required and $\delta$ is called ``online delay''. Notice that $\delta$ is normally a small constant, which is independent of the precision in a given operation. For ease of discussion, in the following of this paper the input data are normalized to fixed point numbers in the range $(-1,1)$. Based on this premise, the online representation of $N$-digit operands and result at iteration $j$ are given by (\ref{Eq:Online_Operands}), where $j\in[-\delta,N-1]$ and $r$ denotes the radix \cite{Ercegovac_Book}.
%
\begin{figure}[tbp]
  \centering
  %\vspace{-2.5ex}
  \includegraphics[width=.48\textwidth]{./Figures/OnlineArithmetic_DataFlow.pdf}
  %\vspace{-3ex}
  \caption{Dataflow in digit-serial online arithmetic, in which both inputs and outputs are processed from the MSD to the LSD. $\delta$ denotes the online delay.}
  %\vspace{-.5ex}
  \label{Fig:OnlineDataFlow}
\end{figure}
%
\begin{eqnarray}\label{Eq:Online_Operands}
\footnotesize
  X_{[j]}=\sum_{i=1}^{j+\delta}x_ir^{-i},~Y_{[j]}=\sum_{i=1}^{j+\delta}y_ir^{-i},~Z_{[j]}=\sum_{i=1}^{j}z_ir^{-i}
\normalsize
\end{eqnarray}

MSD-first operation is possible with the employment of the redundant number system. Normally there are 2 most commonly used redundant number representation: carry-save (CS) \cite{CSadder} and signed-digit (SD) \cite{RedundantNumber}. For instance with SD representation, each digit is represented using a redundant digit set $\{-a, \cdots,-1,0, 1, \cdots, a\}$, where $a\in[r/2,r-1]$. In comparison, the standard non-redundant representation only uses a digit set $\{0,r-1\}$. Thus a standard number corresponds to several possible redundant representations. For example, the two's complement number $0.111$ can be represented in the online form as $0.10\overline{1}$, $0.0\overline{1}1$ and $0.111$ among many other possible representations.

Due to the redundancy, the MSDs of the result can be calculated only based on partial information of the inputs, which is required by digit-serial online arithmetic. Then the value of the number can be revised by the following digits, because each number embodies multiple representations.

\subsection{Binary Online Addition}
Adders serve as a critical building block for arithmetic operations. To perform digit-parallel online addition, a redundant adder can be directly utilized. The structure diagram of an online adder with all signals represented with SD numbers is shown in Fig.~\ref{Fig:Radix2SD_adder}. An major advantage of the redundant number system over the standard ripple-carry based arithmetic is that the propagation of carry is eliminated, resulting in a precision-independent computation time for addition. As labelled in Fig.~\ref{Fig:Radix2SD_adder}, ideally the computation delay of this adder is only 2 full adder (FA) delays for any operand word-length, with the cost of one extra FA for each digit of operands. This makes the online adder suitable for building up more complex arithmetic operators such as multipliers to accelerate the sum of partial products \cite{RedundantMult_1987}.

\begin{figure}[htbp]
	\centering
	\includegraphics[width=.5\textwidth]{./Figures/SDAdder.pdf}
    \vspace{-5ex}
	\caption{An $N$-digit binary digit-parallel online adder with $(N+1)$-digit outputs. Both inputs and outputs are represented using SD representation.}
    \vspace{2ex}
	\label{Fig:Radix2SD_adder}
\end{figure}

@@

\subsection{Binary Online Multiplication}
Multiplication is another key arithmetic operator. Typically the online multiplication is performed in a recursive digit-serial manner, as illustrated in Algorithm~\ref{Algorithm:OnlineMult} \cite{Ercegovac_OnlineMult} where both inputs and outputs are $N$-digit number as given in~(\ref{Eq:Online_Operands}). For a given iteration $j$, the product digit $z_j$ is generated with MSD-first through a selection function $sel()$. For the radix $r$ and a chosen digit set, there exits an appropriate selection method and a value of $\delta$ which ensure convergence \cite{Ercegovac_OnlineMult}. As radix-2 is used most commonly in computer arithmetic, we keep $r=2$ throughout this paper with the corresponding redundant digit set $\{\overline{1},0,1\}$. In this case $sel()$ is given by (\ref{Eq:SelFunc_OM}) \cite{Oregon_OnlineNetwork}. Notice that the election is made only based on 2 integer digits and 1 fractional digit of $W_{[j]}$.

\begin{algorithm}[tbp]
  \caption{Online Multiplication}  \label{Algorithm:OnlineMult}
  \begin{algorithmic}[1]
   \State \textbf{Initialization:}~$X_{[-\delta]}=Y_{[-\delta]}=P_{[-\delta]}=0$\vspace{.5ex}
    \For{$j=-\delta,~-\delta+1,~\cdots,~2N-1$}\vspace{.5ex}
        \State $H_{[j]} \leftarrow r^{-\delta}\left(x_{j+\delta+1}\cdot Y_{[j+1]}+y_{j+\delta+1}\cdot X_{[j]}\right)$\vspace{.5ex}
        \State $W_{[j]} \leftarrow P_{[j]} + H_{[j]}$\vspace{.5ex}
        \State $Z_{[j]}~ \leftarrow sel(W_{[j]})$\vspace{.5ex}
        \State $P_{[j+1]} \leftarrow r\left(W_{[j]}-Z_{[j]}\right)$\vspace{.5ex}
    \EndFor
  \end{algorithmic}
\end{algorithm}

@@

In Section~\ref{Sec:OM_FPGA}, we will first describe an optimized online multiplication algorithm and its FPGA implementation, which are designed specifically targeting on digit parallel operations. Then we provide a modified structure which only generates the most significant half the products, in order to achieve extra area savings.

@@

\begin{enumerate}
    \item efficient mapping of the digit-parallel online adder onto sophisticated FPGAs resources \cite{FPT09RA,ASAP09RA};
    \item multiple operands addition by designing compressor trees based on bit counters \cite{TC13RA,ASPDAC10RA_FPGA};
    \item modifying existing FPGA architecture for more efficient digit-parallel online addition \cite{DAC07_CounterTree} and for specific applications \cite{FPL09_onlineCCM}.
\end{enumerate}

Type 2 and type 3 are beyond the scope of this paper, as our interests are in the field of two-operand online addition that performs $Z=X+Y$, as shown in Fig.~\ref{Fig:Radix2SD_adder}, and we are using the main-stream FPGAs for implementation.

@@

\begin{figure}[tbp]
	\centering
	\includegraphics[width=.5\textwidth]{./Figures/SDAdder_42comp_Spartan.pdf}
	\caption{Map the online adder onto Spartan FPGAs using the fast-carry resources. The grey background highlights the 4:2 compressor. Dotted circle indicates the logic block (LB) which can be mapped to the FPGA using the carry resources.}
	\label{Fig:PreviousWork}
\end{figure}

However, we notice the major limitations of both approaches that they only target on FPGAs with 4-input LUTs (4-LUT) and 2 LUTs within a logic Slice, such as the Xilinx Spartan series and the Altera Cyclone series. This is naturally reasonable because one \texttt{LB} can be mapped to a single Slice. Nevertheless, for FPGAs with 6-input LUTs (6-LUT) and 4 LUTs in a Slice, such as the Xilinx Virtex series and all Xilinx 7 series FPGAs, directly applying these approaches will result in either resource waste or logic fault. For instance, if two \texttt{LB}s are mapped to a slice with 4 LUTs as seen in Fig.~\ref{Fig:PreWorkVirtex6}, the outputs of \texttt{LB1} will be faulty because the its carry input cannot be explicitly initialized. Additionally, the output of \texttt{XORCY2} should not be used.

@@

\subsection{Proposed Mapping Method}
To tackle this problem, we firstly modify the structure of the online adder to enable an efficient FPGA mapping. The structure of a 4-digit online adder is given as an example in Fig.~\ref{Fig:ModifiedOA}. In this equivalent structure, the first FA in each 4:2 compressor is split into 2 parts, which only generate carry and sum respectively. In this case they can be mapped individually on 2 LUTs. The second FA, which generates the outputs, is unchanged and can be implemented using the fast-carry logic.

%propose an alternative mapping method that targets on the slice containing 4 LUTs such as the Xilinx Virtex series and all Xilinx 7 series FPGAs.

\begin{figure*}[tbp]
	\centering
	\includegraphics[width=.85\textwidth]{./Figures/SDadder_FPGA_New_4digit.pdf}
	\caption{Modified structure of online adder. Left: an example of 4-digit online adder. The shaded part refers to the 2 logic blocks (LBs) that can be mapped onto 1 slice. Right: one LB. The dotted box outlines the logic that can be mapped onto 1 LUT and the corresponding fast-carry logic.}
	\label{Fig:ModifiedOA}
\end{figure*}

The detailed slice mapping of the 2 \texttt{LB}s and the logic in each LUT are shown in Fig.~\ref{Fig:SliceNew}. The I/O signals are identical to the previous example in Fig.~\ref{Fig:ModifiedOA}. Normally 6-LUT can be configured with two different output ports \texttt{O6} and \texttt{O5}. For \texttt{LB1}, the carry input can be initialized by setting the \texttt{O6} of LUT2 to 0 constantly. In this case, the output of \texttt{MUXCY2} is always \texttt{O5} of LUT2, and the carry from \texttt{LB2} will not affect the results of \texttt{LB1}. For those logics to generate the MSD and the LSD, they can be combined into 1 Slice as observed from Fig.~\ref{Fig:ModifiedOA}. Therefore using this mapping method the resources within 1 slice can be fully utilized, potentially leads to significant area reduction.

\begin{figure}[tbp]
	\centering
	\includegraphics[width=.5\textwidth]{./Figures/SDadder_FastCarry_New.pdf}
	\caption{Implementation of 2 logic blocks (LBs) in 1 FPGA slice which contains four 6-LUTs. NC stands for ``Not Care''.}
	\label{Fig:SliceNew}
\end{figure}

In this mapping approach, the theoretical area figures of the online adder in terms of the number of LUTs and Slices with respect to the operand word-lengths ($N$) are given in (\ref{Eq:OAareaLUT}) and (\ref{Eq:OAareaSlice}), respectively.
\begin{eqnarray}\label{Eq:OAareaLUT}
	OAarea\_LUT = 2N
\end{eqnarray}
%
\begin{eqnarray}\label{Eq:OAareaSlice}
	OAarea\_Slice = 1+\left\lceil\frac{N-2}{2}\right\rceil
\end{eqnarray}

\subsection{Performance Analysis}\label{subsec:OA_Performance}
In our experiments, we compare the proposed online adder (\texttt{OA\_new}) against the original design which is implemented using behavioural descriptions based on 4:2 compressors (\texttt{OA\_behv}).  In addition, we also include ripple carry adder (\texttt{RCA}) which uses conventional arithmetic, into the comparisons. All designs are targeting on Xilinx Virtex-6 FPGA with speed grade -1. Notice that in order to utilize the carry resources and to avoid logic optimizations by the synthesis tool, the Xilinx primitive component "\texttt{Carry4}" is used to create \texttt{OA\_new}.

We record the rated frequencies of all designs when varying the operand word-lengths. The results are shown in Fig.~\ref{Fig:AdderFreq}. The frequency values are obtained through Xilinx Timing Analyzer after placing and routing the designs in ISE 14.1. It can be seen that both designs with online arithmetic achieves relatively stable operating frequencies across a variety of operand word-lengths. This is as expected, because the critical path delay in the online adder is theoretically a fixed value and it is irrelevant to the operand precisions. In comparison, the rated frequency of \texttt{RCA} drops drastically for larger designs. In this case, both online adders outperform \texttt{RCA} for large word-lengths. Besides, our approach operates constantly faster than \texttt{OA\_behv}. The speed-ups of \texttt{OA\_new} over \texttt{OA\_behv} are from $20\%$ to $26\%$.

@@

We also compare the resource usages of all three adder designs with respect to a variety of operand word-lengths. The results are presented in Fig.\ref{Fig:AdderArea}, in which the theoretical area budgets of the online adder predicted by (\ref{Eq:OAareaLUT}) and (\ref{Eq:OAareaSlice}) are also plotted with dotted lines. Several observations can be made from the results. Firstly the predicted resource usages of both LUTs and Slices match well the results obtained from real implementations. Secondly it can be seen that the behavioural implementation \texttt{OA\_behv} is not area efficient for FPGAs, because the dedicated FPGA resources are not fully utilized. In comparison, our implementation \texttt{OA\_new} achieves significant area savings: $25\%\sim 33\%$ in LUTs and $67\%\sim77\%$ in Slices. Thirdly, our design still costs more area than \texttt{RCA}, with the area overhead of $80\%\sim98\%$ in terms of LUTs, and $0\%\sim88\%$ in terms of Slices. This is because for a single adder, the area overhead brought by the redundant number representations cannot be removed. 
%However research has shown that this overhead can be reduced for multiple-operand adders or adder trees [xxx].

>>>>

\begin{abstract}
%\boldmath
Online arithmetic has been widely utilized for ASIC implementations, as it is designed to perform computations from the most significant digits. Recent research also shows that digit parallel online operators can fail more gracefully when operating beyond the deterministic region, in comparison to operators with conventional arithmetic. Unfortunately, online arithmetic operators require a large area overhead, while efficient implementation of general purpose online operators on FPGAs remains under investigation. In this paper, we propose novel approaches to implement key primitives of online arithmetic, such as adders and multipliers, efficiently on modern FPGAs with 6-input LUTs. The FPGA carry resources are used for both area reduction and performance improvements. We demonstrate experimentally that the proposed architectures achieve significant area savings and frequency speed-ups compared to the original implementations.

@@

\section{Introduction}\label{Sec:Intro}
In the conventional form of computer arithmetic, the computation results are generated either from the least significant digit (LSD), e.g. addition and multiplication, or from the most significant digit (MSD), e.g. division and square root. This inconsistency in computing directions will potentially result in large latency when propagating data among different operations. Online arithmetic was designed to solve this problem \cite{Ercegovac_OnlineOverview,Ercegovac_Book}. With online arithmetic, both inputs and outputs are processed in a MSD-first manner. This enables parallelism among multiple operations, and the overall computation latency can be significantly reduced. A brief overview of online arithmetic is given in Section~\ref{Sec:Background}.

Recent research has demonstrated another key advantage of online arithmetic is that it allows gracefully degradation when operating beyond the deterministic region, e.g. under overclocking \cite{SKDAC14}. This stream of research stems from the inspiration that for certain applications, releasing the absolute accuracy requirements could lead to significant performance improvements. However, traditional forms of computer arithmetic do not fail gracefully when timing violation happens, because timing errors initially affect MSDs of the results. In comparison, online arithmetic is more ``overclocking friendly'', as timing errors only occur at LSDs of the results.

Despite its attractiveness, the usage of online arithmetic on existing FPGAs is still limited due to the huge area overhead. Generally, it is difficult to efficiently map the arithmetic circuits solely onto the look-up-tables (LUTs), which are fundamental building blocks of FPGAs. Lots of efforts has been done to alleviate this concern. For instance, both main commercial FPGA vendors introduced dedicated carry logic into the logic block to reduce the carry propagation delay by over one order of magnitude~\cite{Virtex6}. There is a large amount of studies exploring methods that can effectively map circuits using the carry logic for performance improvements~\cite{FPL10_FPGA_CarryChain,Prefix_CarryChain}, or propose alternative carry logic for FPGAs~\cite{FPGA_CarryChain_New1,FPGA_CarryChain_New2}. Furthermore, hard DSP blocks are also included to boost multiplications and multiply-accumulations~\cite{XilinxDSP}. However, most of these approaches are designed to accelerate computations with the conventional form of arithmetic.

For online arithmetic, we notice from existing literature that there are two directions worth investigating. First, although there has been previous research about FPGA implementations of online arithmetic, they typically only focus on online adders. A brief review of relevant works in this area is presented in Section~\ref{Sec:OA_FPGA}. However, the implementations of other key arithmetic primitives, such as online multipliers, lack exploration. Second, we found that existing approaches of building general purpose online adders are specifically for low cost FPGAs with 4-input LUTs and 2 LUTs within a Slice, such as the Altera Cyclone III \cite{AlteraCyclone} and Xilinx Spartan~3 series \cite{XilinxSpartan}. However, as will be discussed in Section~\ref{Sec:OA_FPGA}, these approaches cannot be directly applied on many modern FPGA with 6-input LUTs and 4 LUTs in a Slice, such as the Xilinx Virtex series FPGAs and all Xilinx 7 series FPGAs~\cite{Virtex7}.

In this paper, both issues will be addressed. In Section~\ref{Sec:OA_FPGA}, we propose a novel approach of mapping the digit parallel online adder onto FPGA with 6-LUTs. The mapping is based on the fast carry logic, and the available resources within a Slice can be fully utilized. In Section~\ref{Sec:OM_FPGA}, we optimize the online multiplication algorithm to yield an efficient FPGA implementation of digit parallel online multiplier. In addition, the theoretical resource usage of both operators is given. We demonstrate experimentally on Xilinx Virtex-6 FPGAs that the proposed designs achieve significant area reduction and performance gain over their original implementations. In summary, the main contributions of this paper are as follows:
\begin{enumerate}
    \item An efficient approach of mapping online adders onto modern FPGAs;
    \item Area efficient implementation of digit parallel online multiplication algorithm on FPGAs;
    \item Theoretical area models of the proposed architectures.
\end{enumerate}

@@

Online arithmetic has been widely used in numerous applications such as signal processing and control algorithms~\cite{Online_FPGADSP,Online_Control}. Online arithmetic was originally designed for digit-serial operation, of which the data flow is illustrated in Fig~\ref{Fig:OnlineDataFlow}. It can be seen that in order to generate the first output digit, $\delta$ digits of inputs are required and $\delta$ is called ``online delay''. Notice that $\delta$ is normally a small constant, which is independent of the precision in a given operation. For ease of discussion, for the rest of this paper, the input data is normalized to fixed point numbers in the range $(-1,1)$. Based on this premise, the online representation of $N$-digit operands and result at iteration $j$ are given by (\ref{Eq:Online_Operands}), where $j\in[-\delta,N-1]$ and $r$ denotes the radix \cite{Ercegovac_Book}.
%
\begin{figure}[tbp]
  \centering
  %\vspace{-2.5ex}
  \includegraphics[width=.48\textwidth]{./Figures/OnlineArithmetic_DataFlow.pdf}
  \vspace{-4ex}
  \caption{Dataflow in digit-serial online arithmetic, in which both inputs and outputs are processed from the MSD to the LSD. $\delta$ denotes the online delay.}
  \vspace{-1ex}
  \label{Fig:OnlineDataFlow}
\end{figure}
%
\begin{eqnarray}\label{Eq:Online_Operands}
\footnotesize
  X_{[j]}=\sum_{i=1}^{j+\delta}x_ir^{-i},~Y_{[j]}=\sum_{i=1}^{j+\delta}y_ir^{-i},~Z_{[j]}=\sum_{i=1}^{j}z_ir^{-i}
\normalsize
\end{eqnarray}

MSD-first operation is possible with the employment of the redundant number system. Normally there are 2 most commonly used redundant number representation: carry-save (CS) \cite{CSadder} and signed-digit (SD) \cite{RedundantNumber}. For instance with SD representation, each digit is represented using a redundant digit set $\{-a, \cdots,-1,0, 1, \cdots, a\}$, where $a\in[r/2,r-1]$. In comparison, the standard non-redundant representation only uses a digit set $\{0,\cdots,r-1\}$. Thus a standard number corresponds to several possible redundant representations. For example, the two's complement number $0.011$ can be represented in the binary online form as $0.1\overline{1}1$, $0.10\overline{1}$ and $0.011$ among many other possible representations.

Due to the redundancy, the MSDs of the result can be calculated using partial information from both inputs. Then the value of the number can be revised using the subsequent digits, because each number embodies multiple representations.

\subsection{Binary Online Addition}
Adders serve as a critical building block for arithmetic operations. To perform digit-parallel online addition, a redundant adder can be used directly. The structure diagram of an online adder with all signals represented with SD numbers is shown in Fig.~\ref{Fig:Radix2SD_adder}. The module 3:2 denotes a 3:2 compressor, which takes 3 inputs and generates 2 outputs. Normally it can be implemented on the basis of a full adder (FA). A major advantage of the redundant number system over the standard ripple-carry based arithmetic is that the propagation of carry is eliminated, resulting in a precision-independent computation time for addition. As labelled in Fig.~\ref{Fig:Radix2SD_adder}, ideally the computation delay of this adder is only 2 FA delays for any operand word-length, with the cost of one extra FA for each digit of operands. This makes the online adder suitable for building up more complex arithmetic operators such as multipliers to accelerate the sum of partial products \cite{RedundantMult_1987}.

\begin{figure}[htbp]
	\centering
	\includegraphics[width=.5\textwidth]{./Figures/SDAdder2.pdf}
    \vspace{-6ex}
	\caption{An $N$-digit binary digit-parallel online adder. Both inputs and outputs are represented using SD representation. Module 3:2 denotes the 3:2 compressor.}
    %\vspace{2ex}
	\label{Fig:Radix2SD_adder}
\end{figure}

@@

\subsection{Binary Online Multiplication}
Multiplication is another key arithmetic operator. Typically, online multiplication is performed in a recursive digit-serial manner, as illustrated in Algorithm~\ref{Algorithm:OnlineMult} \cite{Ercegovac_OnlineMult} where both inputs and outputs are $N$-digit numbers as given in~(\ref{Eq:Online_Operands}). For a given iteration $j$, the product digit $z_j$ is generated MSD-first through a selection function $sel()$. For any radix $r$ and chosen digit set, there exits an appropriate selection method and a value of $\delta$ which ensure convergence \cite{Ercegovac_OnlineMult}. As radix-2 is used most commonly in computer arithmetic, we keep $r=2$ throughout this paper with the corresponding redundant digit set $\{\overline{1},0,1\}$. In this case $sel()$ is given by (\ref{Eq:SelFunc_OM}) \cite{Oregon_OnlineNetwork}. Notice that the selection is based on 2 integer digits and 1 fractional digit of $W_{[j]}$.

\begin{algorithm}[tbp]
  \caption{Online Multiplication}  \label{Algorithm:OnlineMult}
  \begin{algorithmic}[1]
   \State \textbf{Initialization:}~$X_{[-\delta]}=Y_{[-\delta]}=P_{[-\delta]}=0$\vspace{.5ex}
    \For{$j=-\delta,~-\delta+1,~\cdots,~2N-1$}\vspace{.5ex}
        \State $H_{[j]} \leftarrow r^{-\delta}\left(x_{j+\delta+1}\cdot Y_{[j+1]}+y_{j+\delta+1}\cdot X_{[j]}\right)$\vspace{.5ex}
        \State $W_{[j]} \leftarrow P_{[j]} + H_{[j]}$\vspace{.5ex}
        \State $z_{j}~ \leftarrow sel(W_{[j]})$\vspace{.5ex}
        \State $P_{[j+1]} \leftarrow r\left(W_{[j]}-Z_{[j]}\right)$\vspace{.5ex}
    \EndFor
  \end{algorithmic}
\end{algorithm}

@@

In Section~\ref{Sec:OM_FPGA}, we will first describe an optimized online multiplication algorithm and its FPGA implementation, which are designed specifically targeting on digit parallel operations. Then we provide a modified structure which only generates the most significant half of the product in order to achieve extra area savings.

@@

\begin{enumerate}
    \item efficient mapping of the digit-parallel online adder onto sophisticated FPGAs resources \cite{FPT09RA,ASAP09RA};
    \item multiple operands addition by designing compressor trees based on bit counters \cite{TC13RA,CSMult_CounterArray};
    \item modifying existing FPGA architecture for more efficient digit-parallel online addition \cite{DAC07_CounterTree} and for specific applications \cite{FPL09_onlineCCM}.
\end{enumerate}

Type 2 examines the use of counters to sum $k$ $N$-digit numbers where $k>2$. Instead, we focus on two-operand online addition that performs $Z=X+Y$, as shown in Fig.~\ref{Fig:Radix2SD_adder}. In comparison to type 3, this paper attempts to obtain the best performance from mainstream FPGAs instead of suggesting ways to modifying the FPGA fabric.

@@

\begin{figure}[tbp]
	\centering
	\includegraphics[width=.5\textwidth]{./Figures/SDAdder_42comp_Spartan2.pdf}
	\caption{Map the online adder onto Spartan FPGAs using the fast-carry resources. The grey background highlights the 4:2 compressor. Dotted circle indicates the logic block (LB) which can be mapped to the FPGA using the carry resources.}
	\label{Fig:PreviousWork}
\end{figure}

However, we notice the major limitations of both approaches that they only target on FPGAs with 4-input LUTs (4-LUT) and 2 LUTs within a logic Slice, such as the Xilinx Spartan series and the Altera Cyclone series. This is naturally reasonable because one \texttt{LB} can be mapped to a single Slice. Nevertheless, for FPGAs with 6-input LUTs (6-LUT) and 4 LUTs in a Slice, such as the Xilinx Virtex series and all Xilinx 7 series FPGAs, directly applying these approaches will result in either resource waste or logic fault. For instance, if two \texttt{LB}s are mapped to a Slice with 4 LUTs as seen in Fig.~\ref{Fig:PreWorkVirtex6}, the outputs of \texttt{LB1} will be faulty because the its carry input cannot be explicitly initialized. Due to the same reason, the output of \texttt{XORCY2} should not be used.

@@

\subsection{Proposed Mapping Method}
To tackle this problem, we first modify the structure of the online adder to enable an efficient FPGA mapping. The structure of a 4-digit online adder is given as an example in Fig.~\ref{Fig:ModifiedOA}. In this equivalent structure, the first FA in each 4:2 compressor is split into 2 parts, which only generate carry and sum respectively. In this case they can be mapped individually on 2 LUTs. The second FA, which generates the outputs, is unchanged and can be implemented using the fast-carry logic.

%propose an alternative mapping method that targets on the Slice containing 4 LUTs such as the Xilinx Virtex series and all Xilinx 7 series FPGAs.

\begin{figure*}[tbp]
	\centering
	\includegraphics[width=.85\textwidth]{./Figures/SDadder_FPGA_New_4digit2.pdf}
	\caption{Modified structure of online adder. Left: an example of 4-digit online adder. The shaded part refers to the 2 logic blocks (LBs) that can be mapped onto 1 Slice. Right: one LB. The dotted box outlines the logic that can be mapped onto 1 LUT and the corresponding fast-carry logic.}
	\vspace{-1ex}	
	\label{Fig:ModifiedOA}
\end{figure*}

The detailed Slice mapping of the 2 \texttt{LB}s and the logic in each LUT are shown in Fig.~\ref{Fig:SliceNew}. The I/O signals are identical to the previous example in Fig.~\ref{Fig:ModifiedOA}. Normally 6-LUT can be configured with two different output ports \texttt{O6} and \texttt{O5}. For \texttt{LB1}, the carry input can be initialized by setting the \texttt{O6} of LUT2 to 0. In this case, the output of \texttt{MUXCY2} is always \texttt{O5} of LUT2, and the carry from \texttt{LB2} will not affect the result of \texttt{LB1}. For those logics to generate the MSD and the LSD, they can be combined into 1 Slice as observed from Fig.~\ref{Fig:ModifiedOA}. Therefore using this mapping method the resources within 1 Slice can be fully utilized, leading to a significant area reduction.

\begin{figure}[tbp]
	\centering
	\includegraphics[width=.5\textwidth]{./Figures/SDadder_FastCarry_New2.pdf}
	\caption{Implementation of 2 logic blocks (LBs) in 1 FPGA Slice which contains four 6-LUTs. NC stands for ``Not Care''.}
	\vspace{-1ex}	
	\label{Fig:SliceNew}
\end{figure}

In this mapping approach, the area figures of the online adder in terms of the number of LUTs and Slices with respect to the operand word-lengths ($N$) are given in (\ref{Eq:OAareaLUT}) and (\ref{Eq:OAareaSlice}), respectively.
\begin{eqnarray}\label{Eq:OAareaLUT}
	OA\_LUT = 2N
\end{eqnarray}
%
\begin{eqnarray}\label{Eq:OAareaSlice}
	OA\_Slice = 1+\left\lceil\frac{N-2}{2}\right\rceil
\end{eqnarray}

\subsection{Performance Analysis}\label{subsec:OA_Performance}
In our experiments, we compare the proposed online adder (\texttt{OA\_new}) against the original design which is implemented using behavioural descriptions based on 4:2 compressors (\texttt{OA\_behv}).  In addition, we also include ripple carry adder (\texttt{RCA}) which uses conventional arithmetic, into the comparisons. All designs are targeting on Xilinx Virtex-6 FPGA with speed grade -1. Notice that in order to utilize the carry resources and to avoid logic optimizations by the synthesis tool, the Xilinx primitive component ``\texttt{Carry4}" is used to create \texttt{OA\_new}.

We record the rated frequencies of all designs when varying the operand word-lengths. The results are shown in Fig.~\ref{Fig:AdderFreq}. The frequency values are obtained through Xilinx Timing Analyzer after placing and routing the designs in ISE 14.1. It can be seen that both designs with online arithmetic achieves relatively stable operating frequencies across a variety of operand word-lengths. This is as expected, because the critical path delay in the online adder is theoretically a fixed value and it is irrelevant to the operand precisions. In comparison, the rated frequency of the \texttt{RCA} drops drastically for larger word-lengths. It follows that both online adders outperform \texttt{RCA} for large word-lengths. Furthermore, our approach operates faster than \texttt{OA\_behv}; the speed-ups of \texttt{OA\_new} over \texttt{OA\_behv} are from $20\%$ to $26\%$.

@@

We also compare the resource usages of all three adder designs with respect to a variety of operand word-lengths. The results are presented in Fig~\ref{Fig:AdderArea}. We can see that in comparison to the original \texttt{OA\_behv}, our implementation \texttt{OA\_new} achieves significant area savings: $25\%\sim 33\%$ in LUTs and $67\%\sim77\%$ in Slices. This is because the dedicated FPGA resources are fully utilized. However, our design still costs more area than an \texttt{RCA}, with an area overhead of $80\%\sim98\%$ in terms of LUTs, and up to $88\%$ in terms of Slices. This is because for a single adder, the area overhead brought by the redundant number representations cannot be removed. 
%However research has shown that this overhead can be reduced for multiple-operand adders or adder trees [xxx].

\begin{figure}[tbp]
  %\vspace{-2.5ex}
  \centering
  \subfigure[]{
  \begin{minipage}{0.45\textwidth}
    %\centering
    \includegraphics[width=.95\textwidth]{./Figures/Exp/Adder_LUTs3.eps}
    \vspace{1ex}
  \end{minipage}%
  }
  \subfigure[]{
  \begin{minipage}{0.45\textwidth}
    %\centering
    \includegraphics[width=.95\textwidth]{./Figures/Exp/Adder_Slices3.eps}
    \vspace{1ex}
  \end{minipage}
  }
  %\vspace{-3ex}
  \caption{Area comparisons of different binary adder implementations with respect to a variety of operand word-lengths. (a) Usage of LUTs. (b) Usage of Slices.}
  \label{Fig:AdderArea}
  \vspace{-1ex}
\end{figure}
